Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-08-08 src/tex/ms.tex
Change for A.2
Changes for A.6

In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
To first order, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
\begin{equation}
    h_{L/R}^{\mathrm{br}}(f) =
    h_{L/R}^{\mathrm{GR}}(f) \times
    \exp\left(\pm\kappa d_C \frac{f}{100\,\mathrm{Hz}}\right)\,,
    \label{eq:waveform_modification}
\end{equation}
where the emitted waveform $h_{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_{L/R}^{\mathrm{br}}$.
The overall magnitude of this effect for a given frequency $f$ is set by an attenuation coefficient, $\kappa$, which encodes the intrinsic strength of the birefringence.
The emitted waveform for a given source (i.e., the waveform observed in the near zone, very close to the source) will generally differ from the analogous waveform predicted by \ac{GR} \cite{Alexander:2009tp,Okounkova:2019zjf}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small (e.g., \cite{Okounkova:2022grv}), it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).

@@

\subsubsection{Wang et al.}

@@

The parameterization of \citet{Wang_2021} is the same as that in the more recent work by \citet{Zhu:2023wci}.
That work, which appeared while this manuscript was finalized, reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}$ based on GWTC-3.

>>>>

In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening.
Similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci}.
To first order, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
\begin{equation}
    h_{L/R}^{\mathrm{br}}(f) =
    h_{L/R}^{\mathrm{GR}}(f) \times
    \exp\left(\pm\kappa d_C \frac{f}{100\,\mathrm{Hz}}\right)\,,
    \label{eq:waveform_modification}
\end{equation}
where the emitted waveform $h_{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_{L/R}^{\mathrm{br}}$.
The overall magnitude of this effect for a given frequency $f$ is set by an attenuation coefficient, $\kappa$, which encodes the intrinsic strength of the birefringence.
The emitted waveform for a given source (i.e., the waveform observed in the near zone, very close to the source) will generally differ from the analogous waveform predicted by \ac{GR} \cite{Alexander:2009tp,Okounkova:2019zjf}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small (e.g., \cite{Okounkova:2022grv}), it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).

@@

\subsubsection{Wang et al.}
\label{sec:comparison_Wang}

@@

\subsubsection{Zhu et al.}
A more recent work by \citet{Zhu:2023wci}, which appeared while this manuscript was finalized, performed \ac{PE} on GWTC-3 events with a frequency-dependent model of amplitude birefringence.
Their model is similar to the one used in \citet{Wang_2021}, which allows us to compare our results, as we did in Sec.~\ref{sec:comparison_Wang}.
\citet{Zhu:2023wci} reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Zhu.txt}.
This result constrain the birefringence effect much weaker than our result.
This is an unexpected result, as the analysis in \citet{Zhu:2023wci} is similar to ours\footnote{Note that the prior on the birefringence parameter and luminosity distance are different between the analysis by \citet{Zhu:2023wci}and ours, but it is unlikely to be the cause of the difference in the constraint.} and they included more events than we did.
However, we could not compare the analysis in \citet{Zhu:2023wci} with ours in detail, as the analysis is not publicly available, and the list of events included is not provided in the paper.
